---
title: "Contributing to qData"
author: "James Hollway"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Contributing to qData}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Contributing to the qData ecosystem of packages

`{qData}` has package related functions that help developers contribute to the ecosystem of 
packages by helping with the set-up of packages consistent with `{qData}`. 
The function `setup_package()`, for example, establishes many of the required files and folder 
structures for a `{qData}`-consistent data package. 
It requires that the package name starts with a the letter "q" as well as that 
the name of at least one author is declared. For author names, please declare surname(s) first and then given name(s) separated by commas.
The function accepts ORCID numbers instead of author names as well. 

```{r eval=FALSE}
setup_package("qStates", AuthorName = c("Hollway, James", "Sposito, Henrique"), Role = list(c("cre", "aut", "ctb"), c("ctb")))
```

For adding on package contributors or authors to q Package later on, use to the `new_author()` function.
This function also accepts ORCID numbers instead of author names. 

```{r eval=FALSE}
new_author(AuthorName = c("Bieri, Bernhard", "Peev, Esther"), Role = c("ctb", "ctb"))
```

## Collecting data

For data collection and loading into a new package with `{qData}`, the function `import_data()` 
creates a data-raw folder into the package, copies the raw data into the folder and 
provides a script to facilitate data loading and data correction using `{qData}`. 
The function requires that a name for the dataset - two dimensional data sheet being 
imported - is declared, in upper case letters, as well as the name of the issue-domain, 
in lower case letters, the dataset belongs. The domain issue name will be used to connect 
two-dimensional datasets into three-dimensional databases that resembles a data cube (a list object where entries are tibbles).

```{r eval=FALSE}
qData::import_data("COW", "states")
```

## Correcting data

Once a dataset is collected with `import_data()`, it needs to be corrected so that it can be 
connected into a database. `{qData}` contains several functions geared towards data wrangling. 
These include, for example, the `transmutate()` function which is between `{dplyr}`'s 
transmute and mutate but that returns only the mutated variables and none of the variables
used in the mutations. `{qData}` also includes functions that help correct and standardise dates. 
These include, for example, the `standardise_dates()` function which creates nested vectors of 
dates for vague date inputs into a range of dates. As well, the `standardise_title()` function 
capitalises all words in a string and removes unnecessary symbols to enable comparison.

```{r eval=FALSE}
COW <- transmutate(ID = stateabb,
                   Beg = standardise_dates(styear, stmonth, stday),
                   End = standardise_dates(endyear, endmonth, endday),
                   Label = standardise_titles(statenme),
                   COW_Nr = standardise_titles(as.character(ccode)))
```

Contributors should make sure the variables names in the datasets are in line with `{qData}` ecosystem.
Datasets part of the same database should have similar variables which allows to build up the 'datacube' 
for comparison.

## Connecting data

Once the dataset has been corrected in `{qData}`, the cleaned dataset can be loaded into the 
package using the `export_data()` function which take as inputs the dataset name and the database name as a string. 
Corrected datasets can be connected to form a three dimensional database that resembles a data cube.
The website source of the data should be indicated in the third argument of the function.
The function creates a database in a data folder to store corrected datasets, while providing a citation template and a test template.
The citation template helps cite the source of the raw data and and describe the variables in the corrected dataset. 
The test template makes sure the corrected dataset is consistent with `{qData}` guidelines.

```{r eval=FALSE}
qData::export_data(COW, "states", URL = "https://correlatesofwar.org/data-sets/state-system-membership")
```

The test template is meant to ensure consistency for the corrected dataset. 
If these tests fail, users can go back to correcting the data and, then, re-run `export_data()`.  
