---
title: "Contributing qData"
author: "James Hollway"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Contributing qData}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Contributing to the qData ecosystem of packages

qData has package related functions that help developers contribute to the ecosystem of packages by helping with the set-up of packages 
consistent with qData. 
The function `setup_package()`, for example, establishes many of the required files and folder 
structures for a qData-consistent data package. 
It requires that the package name starts with a the letter q as well as that 
the name of at least one author is declared. 

```{r}
qData::setup_package("qStates", "James Hollway")
```

## Collecting data 

For data collection and loading into a new package with qData, the function `import_data()` creates a data-raw folder into the package, 
copies the raw data into the folder and provides a script to facilitate data loading and data correction using qData. 
The function requires that a name for the dataset - two dimensional data sheet being imported - is declared, in upper case letters, 
as well as the name of the issue-domain, in lower case letters, the dataset belongs. 
The domain issue name will be used to connect two-dimensional datasets into three-dimensional databases that resembles a data cube.

```{r}
qData::import_data("COW", 'states')
```

## Correcting data

Once a dataset is collected with `import_data()`, it needs to be corrected so that it can be connected into a database. 
qData contains several functions geared towards data wrangling. 
These include, for example, the `transmutate ()` function which is between between dplyr's transmute and mutate but that returns 
only the mutated variables and none of the variables used in the mutations. 
Other useful functions for correcting data with qData are geared towards data standardising. 
These include, for example, the `entitle()` function that capitalises all words in a string to enable comparison.

```{r}
COW <- transmutate(statename = entitle(statenme)) %>% 
transmutate(Id = stateabb) %>% 
transmutate(stdate = paste(stday, stmonth, styear, sep = ".")) %>% 
transmutate(edate = paste(endday, endmonth, endyear, sep = ".")) %>% 
transmutate(Beg = lubridate::dmy(stdate)) %>% 
transmutate(End = lubridate::dmy(edate))  
```

qData also includes functions that help correct and standardise dates. These include, for example, the `standardise_dates()` 
function which creates nested vectors of dates for vague date inputs into a range of dates. 

```{r}
qData::standardise_dates(COW)
```

Once the dataset has been corrected in qData, the cleaned dataset can be loaded into the package using the `export_data()` function. 
The function creates a data folder to store corrected datasets, while providing a citation template and a test template. 
The citation template helps cite the source of the raw data and and describe the variables in the corrected dataset. 
The test template makes sure the corrected dataset is consistent with qData guidelines.

```{r}
qData::export_data(COW)
```

The test template is meant to ensure consistency for the corrected dataset. 
If these tests fail, users can go back to correcting the data and, then, re-run `export_data()`.  

## Connecting data

Corrected datasets can be connected to form a three dimensional database that resembles a data cube. 
This can be done with the help of several functions in qData. 
