---
title: "Working with qData"
author: "James Hollway"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working with qData}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Discovering data

The first thing users of the package will want to do is to identify
datasets that might contribute to their research goals. Since some of
these data packages are too big for CRAN (the main though not only R
package repository), we expect that their developers will instead choose
to make their packages available on Github. Github is huge though,
either way it can be difficult to identify salient data. We have tried
to make this easy by allowing the reporting of all datasets in the
ecosystem that we know of that currently pass our tests.

```{r}
library(qData)
get_packages()
```

In the future, this function will report more details about the datasets
included in each package and their provenance.

For now, let us say that we wish to download the `{qStates}` package,
which offers a set of datasets related to state actors in global
governance. We can download and install the latest release version of
the `{qStates}` package using the same function as before, only
specifying which package we want to 'get':

```{r}
get_packages("globalgov/qStates")
```

## Loading data

Once the qpackage has been downloaded and installed, we are free to work
with the data it contains. This `{qStates}` package includes several
datasets. We can get a quick summary of the datasets included in this
package with the following command:

```{r}
data(package = "qStates")
```

Let's take a quick look at the main `states` data in this package.

```{r}
states <- qStates::states 
# You can also get this like so: 
# data(states, package = "qStates")
states
```

We can see that there are three named datasets relating to states here:
`COW` (Correlates of War), `GW` (Gleditsch and Ward), and `ISD`
(International Systems Dataset). Each of these datasets has their
advantages and so we may wish to understand their differences, summarise
variables across them, and perhaps also rerun models across them.

To retrieve an individual dataset from this database,
we can use the `pluck()` function from the `{purrr}` package.

```{r}
COW <- purrr::pluck(states, "COW")
```

However, the real value of qData packages is that multiple datasets
relating to the same phenomenon are presented together.
This facilitates comparison, the subject of the next section.

## Comparing data

First of all, we want to understand what the differences are between these datasets.
One important way to understand the relationship between these datasets is 
to understand what their relative advantages and disadvantages are.
For example, one dataset may be long (has many observations) 
while another is shorter but wider (has more variables).
One might include details further back in history or is more recent,
but includes more missing data or less precise data (coded at a less granular level)
than another with a more restrictive 
Or one might appear complete yet offer less information on where the original datapoints
were sourced or how certain variables were coded,
while another provides an extensive and transparent codebook that facilitates replication.
Below we show how one might compare the data in `states` in these ways.

### Differences in dimension

One way to see a quick summary of each of these datasets 
is using the `glimpse()` command from the `{tibble}` package.
`{tibble}` is currently loaded along with `{qData}` and so there is no need to load it independently.

```{r}
glimpse(states)
```

We can see here, for example, that COW seems to have more observations of state sovereignty episodes
than GW does.

### Differences in depth

We can also check whether one dataset has more missings than another.
Here we are going to 'map' a function from the `{Amelia}` package over each of the datasets in the database.
That function is the `missmap()` function.
We will use the `map()` function to do this (though other ways are possible).

```{r, message=FALSE, warning=FALSE}
library(Amelia)
purrr::map(states, missmap)

library(mice)
purrr::map(states, md.pattern, rotate.names = T)
```

### Differences in documentation

We can bring up their documentation using the following command: `?states`.
This documentation outlines where this data was sourced from (when) and what the nature of the main variables are.

## Modelling data

Next we may be interested in whether any relationships we are interested in or inferences we want to draw
are sensitive to which data we use.
That is, we are interested in the robustness of any results to different data specifications.
None of the `states` datasets are particularly wide, but let us try and examine the "Beg" variable to start with.

We can start by exploring whether our conclusion about when states first gained sovereignty
would differ depending on which dataset we use.
We can use the `purrr::map()` function used above, but this time pass it the `mean()` function
and tell it to operate on just the "Beg" variable (removing any NAs).
Since qData datasets are always ordered by "Beg" (and then "ID"),
we can remove any subsequent (duplicated) entries by ID to concentrate on first appearances.

```{r}
states %>% 
  purrr::map(function(x){
    x %>% dplyr::filter(!duplicated(ID)) %>%
      dplyr::summarise(mean(Beg, na.rm = TRUE))
  })
```

It seems that there is some variation there.
Of course, these averages are biased in the sense that these datasets are left-censored.
Any state gaining sovereignty prior to 1816 is simply assigned as "1816-01-01".
