---
title: "Working with qData"
author: "James Hollway"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working with qData}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Discovering data

The first thing users of the package will want to do is to identify datasets that might contribute to their research goals.
Since some of these data packages are too big for CRAN (the main though not only R package repository),
we expect that their developers will instead choose to make their packages available on Github.
Github is huge though, either way it can be difficult to identify salient data.
We have tried to make this easy by allowing the reporting of all datasets in the ecosystem that we know of that currently pass our tests.

```{r}
library(qData)
get_packages()
```

In the future, this function will report more details about the datasets included in each package
and their provenance.

For now, let us say that we wish to download the `{qStates}` package,
which offers a set of datasets related to state actors in global governance.
We can download and install the latest release version of the `{qStates}` package 
using the same function as before, only specifying which package we want to 'get':

```{r, eval=FALSE}
get_packages("qStates")
```

