---
title: "Working with qData"
author: "James Hollway"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working with qData}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Discovering data

The first thing users of the package will want to do is to identify
datasets that might contribute to their research goals. Since some of
these data packages are too big for CRAN (the main though not only R
package repository), we expect that their developers will instead choose
to make their packages available on Github. Github is huge though,
either way it can be difficult to identify salient data. We have tried
to make this easy by allowing the reporting of all datasets in the
ecosystem that we know of that currently pass our tests.

```{r}
library(qData)
get_packages()
```

In the future, this function will report more details about the datasets
included in each package and their provenance.

For now, let us say that we wish to download the `{qStates}` package,
which offers a set of datasets related to state actors in global
governance. We can download and install the latest release version of
the `{qStates}` package using the same function as before, only
specifying which package we want to 'get':

```{r, eval=FALSE}
get_packages("qStates")
```

## Loading data

Once the qpackage has been downloaded and installed, we are free to work
with the data it contains. This `{qStates}` package includes several
datasets. We can get a quick summary of the datasets included in this
package with the following command:

```{r}
data(package = "qStates")
```

Let's take a quick look at the main `states` data in this package.

```{r}
states <- qStates::states 
# You can also get this like so: 
# data(states, package = "qStates")
states
```

We can see that there are three named datasets relating to states here:
`COW` (Correlates of War), `GW` (Gleditsch and Ward), and `ISD`
(International Systems Dataset). Each of these datasets has their
advantages and so we may wish to understand their differences, summarise
variables across them, and perhaps also rerun models across them.

To retrieve an individual dataset from this database,
we can use the `pluck()` function from the `{purrr}` package.

```{r}
COW <- purrr::pluck(states, "COW")
```

However, the real value of qData packages is that multiple datasets
relating to the same phenomenon are presented together.
This facilitates comparison, the subject of the next section.

